\LARGE \bf Explicit Knowledge and the Deliberative Layer: Lessons Learned

### 2013-03-04 paper.tex
Written part 'Architect view' of the conclusion

@@

>>>>

We conclude this article we three different perspective on the knowledge-centric
techniques we have presented here: first, the \emph{architect perspective}:
what are the pro and cons of our approach from an engineer point of view. Then,
the \emph{logician perspective}: does the promise of automated reasoning for
robots really comes to fruition with ontologies? And eventually the
\emph{cognitician perspective} or how explicit knowledge manipulation endows
our robots with new cognitive capabilities.

@@

In this article, we have presented several sub-systems of our robots: a module for geometric
reasoning and situation assessment (written in C++), a natural
language interface (written in Python), a symbolic task planner (written in
C++), two execution controller (one written in Python, one in PRS). These
modules communicate together through RDF statements with only a few exceptions.

Since these statements are anchored in the robot's pool of knowledge, they
actually convey \emph{meaning} through the system (for instance, when the
dialog module or the task planner or the execution controller uses the symbol
{\tt Give} \emph{means}, it refers to the abstract class of all give actions, which are also stated
to be a kind of purposeful actions). This has one important consequence from
the perspective of the system architecture: modules interfaces are only defined
in term of knowledge consumption and/or production. This brings good decoupling
properties. For instance, the output of the dialog module is a new
\emph{situation desired} by the speaker. This module can be transparently
replaced or completed by any other module that produces the same chuncks of
knowledge. Only a thin protocol layer remains to get the modules to interact
together. This allows us to implement (otherwise difficult) merging of
radically different perception modalities like vision (the robot sees an object on
a table) and verbal description (a human says to the robot another object is
under the table): both are converted to RDF statements ({\tt obj1 type Object},
{\tt obj1 isOn table1} and {\tt obj2 type Object}, {\tt obj2 isUnder table1})
and made available to the other robot modules.

Another consequence is that the communication channels are {\it
defacto} defined by the semantics of the information they
convey. This allows the system designers to think in term of which module
produces or needs which knowledge, instead of which module produces which
service. We found this approach to be very helpful when implementing cognitive
functions like the theory of mind, the perspective taking or the natural
language grounding.

Lastly, another good property of knowledge-based communication between module
is the \emph{cognitive observability} of the system: by logging queries between
modules (which is even easier in our case since we rely on a central knowledge
management), one can trace the interactions between the robot systems at a high
abstraction level. We will come back below on the significance of this property
from the cognitive point of view.

The knowledge-centric architecture also has downsides. We can name here a few.

Because the sources of knowledge the robot has to deal with are multiple and
often not known at startup (typically, the knowledge generated by a verbal
interaction with a human), it is difficult to formally guarantee reliability.

Also, it must be noted that actuation itself is directly piloted by the
execution controller through standard RPC calls (ROS actions and/or Pocolibs
requests): no knowledge abstraction takes place.

And obviously, many aspects like uncertainty management, modal logics,
non-monotonic reasoning,... remain to be explored to gain the level of
expressiveness required to cover many of the more complex human-robot
interaction scenarii.
